Place of death in NRS COVID-19 deaths: visualisation for 2022 data

### 2022-02-18 covid_nrs_place_of_death_data_2022.Rmd
changed author name & email

```{r}
weekly_deaths_la <- 
  read_excel(files_weekly_deaths$la, sheet = 2, skip = 2) %>% 
  clean_names %>%
  rename(week_number = week_of_occurrence, ref_area = council, place_of_death = location_of_death, number_of_deaths = deaths) %>%
  filter(!is.na(number_of_deaths)) %>%  # remove blank rows & copyright entry
  parse_year_and_week_number_from_ad_hoc_data %>%
  recode_cause_of_death() %>%
  recalculate_number_of_deaths_after_recoding_cause_of_death %>%  # recompute the sum of deaths
  tidyr::complete(
      nesting(year, week_number, week_number_run_over, date_w_c),  # these are 'identifying' variables that shouldn't be combinatorily exploded
      ref_area, cause_of_death, place_of_death,  # these are the variables we want to have all combinations of, and missing combinations are set to 0 deaths
      fill = list(number_of_deaths = 0)
      ) %>%
  recode_place_of_death() %>%  # this needs to be done before recoding cause of death, otherwise we also include the "all" figure which isn't included in this dataset
  pivot_wider(names_from = cause_of_death, values_from = number_of_deaths) %>%   # make separate columns for causes of death figures
  mutate(deaths_all_causes = deaths_non_covid + deaths_covid_related) %>%  # compute all cause deaths from the other two figures reported
  set_covid_and_non_covid_deaths_to_zero_before_reporting_started()
              quiet = FALSE)

@@

```{r}
historical_average_cause_of_death_2021_onwards <-
  map_dfr(
    .x = cell_ranges_for_causes_of_death_data,
    .f = ~readxl::read_excel(files_weekly_deaths$main_report, sheet = sheet_cause_of_death_data, range = .x), 
    .id = "place_of_death"
  ) %>%
  janitor::clean_names() %>%
  rename(year = registration_year, date_w_c = week_beginning) %>%
  select(place_of_death, year, week_number, date_w_c, matches("five_year_average")) %>%
  pivot_longer(cols = 5:ncol(.), names_to = "name", values_to = "deaths_mean") %>%
  mutate(
    year = as.integer(year),
    cause_of_death = str_remove_all(name, "_five_year_average|_deaths"),
    cause_of_death = return_pretty_cause_of_death(cause_of_death),
    cause_of_death = if_else(cause_of_death == "All causes", "All", cause_of_death),
    date_w_c = as.Date(date_w_c),
    week_number_run_over = compute_week_run_over(week_number, year, start_year = 2020),  # invent a "run-over" week number - weeks 54+ are in 2021, weeks 106+ are in 2022
    historic_period = historic_period_depending_on_year(year)
  ) %>%
  select(-name) %>%
  drop_na(deaths_mean)  # remove blank entries - these are for the current year but for dates in the future
```

@@

merged_deaths_cause_of_death <-
  weekly_deaths_cause_of_death %>%
  mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
  inner_join(
    historical_average_cause_of_death, by = c("place_of_death", "week_number", "cause_of_death", "historic_period", "year", "date_w_c")
  )
```

@@

```{r}
# TODO: work out what to do about this - is it even used?

## bespoke computation because format is different... deaths split by cause!
## Note: this isn't appropriate for 2022 yet because we don't have the annual death figure yet
total_deaths_cause_of_death <-
  merged_deaths_cause_of_death %>%
  group_by(year, place_of_death, cause_of_death, historic_period) %>%
  summarise(deaths = sum(number_of_deaths), .groups="drop") %>%
  left_join(historical_annual_average_cause_of_death, by = c("place_of_death", "cause_of_death", "historic_period")) %>%
  mutate(
    ratio_annual_deaths_to_historical_mean = deaths / annual_deaths_mean
  ) %>%
  filter(year != 2022)

current_and_historical_deaths_by_cause_and_location <-
  bind_rows(
    ## 2020 & 2021 separately
    total_deaths_cause_of_death %>%
      select(year, place_of_death, cause_of_death, deaths) %>%
      mutate(year = as.character(year)),
    ## 2020 & 2021 averaged
    total_deaths_cause_of_death %>%
      select(year, place_of_death, cause_of_death, deaths) %>%
      group_by(place_of_death,  cause_of_death) %>%
      summarise(deaths = sum(deaths), year = "2020-2021", .groups = "drop"),
    ## historical data
    total_deaths_cause_of_death %>%
      select(year, place_of_death, cause_of_death, deaths = annual_deaths_mean) %>%
      mutate(year = "2015-2019") %>%
      distinct
  ) %>%
  filter(place_of_death != "All") %>%
  group_by(year, cause_of_death) %>%
  mutate(proportion = deaths / sum(deaths)) %>%
  ungroup %>%
  arrange(year, cause_of_death)
```

@@

>>>>

```{r}
weekly_deaths_la <- 
  read_excel(files_weekly_deaths$la, sheet = 2, skip = 2) %>% 
  clean_names %>%
  rename(week_number = week_of_occurrence, ref_area = council, place_of_death = location_of_death, number_of_deaths = deaths) %>%
  filter(!is.na(number_of_deaths)) %>%  # remove blank rows & copyright entry
  parse_year_and_week_number_from_ad_hoc_data %>%
  recode_cause_of_death() %>%
  recalculate_number_of_deaths_after_recoding_cause_of_death %>%  # recompute the sum of deaths
  tidyr::complete(
      nesting(year, week_number, week_number_run_over, date_w_c),  # these are 'identifying' variables that shouldn't be combinatorily exploded
      ref_area, cause_of_death, place_of_death,  # these are the variables we want to have all combinations of, and missing combinations are set to 0 deaths
      fill = list(number_of_deaths = 0)
      ) %>%
  recode_place_of_death() %>%  # this needs to be done before recoding cause of death, otherwise we also include the "all" figure which isn't included in this dataset
  pivot_wider(names_from = cause_of_death, values_from = number_of_deaths) %>%   # make separate columns for causes of death figures
  mutate(deaths_all_causes = deaths_non_covid + deaths_covid_related) %>%  # compute all cause deaths from the other two figures reported
  set_covid_and_non_covid_deaths_to_zero_before_reporting_started()

@@

In the `2015-2019` period there were no Covid-19 deaths, so the historical average should be `NA`, which can possibly be interpreted as 0 for the purposes of comparing a year to the historical average.
For the `2016-2019, 2021` period the average number of covid deaths should be equivalent to the deaths in 2021 - we wouldn't want to make an average between 4 years of 0 deaths and one year with any deaths! Similarly if the period were `2017-2019, 2021-2022`, we would compute the average for 2021-2022, and so on.

```{r}
## this is an ugly hack - we set average weekly deaths due to C19 to NA for 2021
## and we set the average covid deaths for 2015-2019, 2021 to the same as in 2021
## however, for the latter, we set the dates to 2022
average_weekly_covid_deaths_2021_onwards <-
  bind_rows(
    ## set 2021 averages to NA
    weekly_cause_of_death_2021_onwards %>%
    filter(cause_of_death == "COVID-19" & year == 2021) %>%
    mutate(number_of_deaths = NA_real_) %>%
    rename(deaths_mean = number_of_deaths),
    ## use 2021 weekly deaths as 2022 historic average
    weekly_cause_of_death_2021_onwards %>%
    filter(cause_of_death == "COVID-19" & year == 2021) %>%
    mutate(year = 2022) %>%
    select(-c(week_number_run_over, date_w_c)) %>%
    rename(deaths_mean = number_of_deaths) %>%
    inner_join(weekly_cause_of_death_2021_onwards %>% 
                select(year, week_number, week_number_run_over, date_w_c, place_of_death, cause_of_death)
              )
  ) %>%
  mutate(historic_period = historic_period_depending_on_year(year))

historical_average_cause_of_death_2021_onwards <-
  bind_rows(
    ## the hacked averages for 2021 and 2022 above
    average_weekly_covid_deaths_2021_onwards,
    ## extract averages from most recent file for 2021 onwards
    map_dfr(
      .x = cell_ranges_for_causes_of_death_data,
      .f = ~readxl::read_excel(files_weekly_deaths$main_report, sheet = sheet_cause_of_death_data, range = .x), 
      .id = "place_of_death"
    ) %>%
    janitor::clean_names() %>%
    rename(year = registration_year, date_w_c = week_beginning) %>%
    select(place_of_death, year, week_number, date_w_c, matches("five_year_average")) %>%
    pivot_longer(cols = 5:ncol(.), names_to = "name", values_to = "deaths_mean") %>%
    mutate(
      year = as.integer(year),
      cause_of_death = str_remove_all(name, "_five_year_average|_deaths"),
      cause_of_death = return_pretty_cause_of_death(cause_of_death),
      cause_of_death = if_else(cause_of_death == "All causes", "All", cause_of_death),
      date_w_c = as.Date(date_w_c),
      week_number_run_over = compute_week_run_over(week_number, year, start_year = 2020),  # invent a "run-over" week number - weeks 54+ are in 2021, weeks 106+ are in 2022
      historic_period = historic_period_depending_on_year(year)
    ) %>%
    select(-name) %>%
    drop_na(deaths_mean)  # remove blank entries - these are for the current year but for dates in the future
  )
```

@@

merged_deaths_cause_of_death <-
  weekly_deaths_cause_of_death %>%
  mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
  left_join(  # left_join so that the historical Covid-19 rates are blank!
    historical_average_cause_of_death %>% select(-week_number_run_over), 
    by = c("place_of_death", "week_number", "cause_of_death", "historic_period", "year", "date_w_c")
  )
```

@@

```{r}
## bespoke computation because format is different... deaths split by cause!
total_deaths_cause_of_death <-
  merged_deaths_cause_of_death %>%
  group_by(year, place_of_death, cause_of_death, historic_period) %>%
  summarise(deaths = sum(number_of_deaths, na.rm = TRUE), .groups="drop") %>%
  left_join(historical_annual_average_cause_of_death, by = c("place_of_death", "cause_of_death", "historic_period")) %>%
  mutate(
    ratio_annual_deaths_to_historical_mean = if_else(  # only compute this for complete years, we can't compare 2022 to annual average!
      condition = year == 2022,
      true = NA_real_,
      false = deaths / annual_deaths_mean
    )
  )

current_and_historical_deaths_by_cause_and_location <-
  bind_rows(
    ## 2020-2022 separately
    total_deaths_cause_of_death %>%
      select(year, place_of_death, cause_of_death, deaths) %>%
      mutate(year = as.character(year)),
    ## historical data
    total_deaths_cause_of_death %>%  # these are for 2020-2021, but we use the historical mean
      filter(year == 2020) %>%
      select(year, place_of_death, cause_of_death, deaths = annual_deaths_mean) %>%
      mutate(year = "2015-2019")
  ) %>%
  filter(place_of_death != "All") %>%
  group_by(year, cause_of_death) %>%
  mutate(proportion = deaths / sum(deaths)) %>%
  ungroup %>%
  arrange(year, cause_of_death)
```

@@

# Annual deaths by place, 2015-2022

```{r}
merged_proportions_of_deaths_by_place <-
  bind_rows(
    ## 2015-2019
    weekly_deaths_2015_2019_overall %>% 
    group_by(year, place_of_death) %>% summarise(number_of_deaths = sum(number_of_deaths), .groups="drop") %>%
    filter(place_of_death != "All" & !is.na(as.integer(year))) %>%
    group_by(year) %>%
    mutate(
      total_deaths = sum(number_of_deaths)
    ) %>%
    ungroup %>%
    mutate(
      proportion_of_total = number_of_deaths / total_deaths, 
      year=as.integer(year)
      ),
    ## 2020 onwards
    total_deaths_overall %>%
    select(year, place_of_death, number_of_deaths = deaths_all_causes) %>%
    distinct %>%  # remove duplicate entries for 2022 due to having multiple possible historic periods
    filter(place_of_death != "All") %>%
    group_by(year) %>%
    mutate(
      total_deaths = sum(number_of_deaths)
    ) %>%
    ungroup %>%
    mutate(proportion_of_total = number_of_deaths / total_deaths)
  )
```

### 2022-02-18 covid_nrs_place_of_death_visualisations_2022.Rmd
changed author name & email

Note: visualisations have been moved to a separate .Rmd file

@@

date_range_text <- glue::glue("Data from w/c {format(earliest_date, '%d %B %Y')} to w/c {format(most_recent_date_available_2021, '%d %B %Y')} (inclusive).")

@@

save_output_file(filename = file.path(dir_outputs, paste0(c("fig_place_of_death", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = figure_place_of_death, width = 28, height = 14, dpi = 300, units = "cm")
```

@@

```{r}
figure_home_deaths_over_time_by_hb_4_week_average <-
  merged_deaths_hb %>%
  filter(place_of_death == "Home & other non-institution") %>%
  mutate(ref_area = factor(x = ref_area, levels = order_hb_by_population)) %>%  # arrange hb/la by population size
  pivot_longer(cols = c(ma4w_deaths_all_causes, ma4w_deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2022)) +
  geom_ribbon(aes(ymin = ma4w_deaths_min, ymax = ma4w_deaths_max), fill = "#4477aa", alpha = 0.5) +
  # geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = ma4w_deaths_mean), linetype = "dashed", size = 0.2) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  # scale_x_continuous(breaks = seq(0,60,by=10)) +
  facet_wrap(~ref_area, scales = "free_y") +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  theme(
    legend.position = "top", 
    plot.caption = element_text(size = 10, colour = "gray60"),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
    ) +
  labs(x = NULL,
       y = "Deaths (average of past 4 weeks)",
       title = "Deaths at Home*: 4-week average by NHS Health Board",
       subtitle = date_range_text,
       caption = paste0(c(
         "*Home or other non-institutional setting", vertical_line_text, vertical_scale_differs_between_panels_text, source_nrs_text
         ), collapse="\n"),
       colour = "Cause of death"
       ) +
  ## annotate historic range
  annotate(geom = "text", x=ymd("2021-01-07"), y=0, label = "2021", hjust = 0, vjust = -1, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=0, label = "'22", hjust = 0, vjust = -1, colour = "grey40") +
  NULL

save_output_file(filename = file.path(dir_outputs, paste0(c("fig_home_deaths_4week_average_over_time_by_hb", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = figure_home_deaths_over_time_by_hb_4_week_average, width = 30, height = 20, dpi = 300, units = "cm")
```

>>>>

@@

date_range_text <- glue::glue("Data from w/c {format(earliest_date, '%d %B %Y')} to w/c {format(most_recent_date_available_2022, '%d %B %Y')} (inclusive).")

@@

figure_place_of_death %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_place_of_death", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 28, height = 14, dpi = 300, units = "cm")
```

## Home deaths only

```{r}
figure_place_of_death_home_only <-
  merged_deaths_overall %>%
  filter(sex == "all" & age == "all") %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%  # home death only
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2022)) +
  geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = deaths_mean), linetype = "dashed", colour="grey20") +
  facet_grid(~place_of_death) +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = NULL,
       y = "Deaths (N)",
       subtitle = date_range_text,
       caption = paste0(c(death_registrations_text,removed_other_category_text,source_nrs_text), collapse="\n")
       ) +
  ## annotate historic range
  annotate(geom="text",x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 50, label = "Historic range (min & max)", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 29, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 33, year_number = 2020), y = 65, yend = 210, curvature = 0.20, colour = "#4477aa", alpha = 0.5, arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom="text", x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 100, label = "Historic mean", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 22, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 25, year_number = 2020), y = 115, yend = 270, curvature = 0.20, colour = "grey20", linetype="dashed", arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom = "text", x=ymd("2021-01-07"), y=30, label = "2021", hjust = 0, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=30, label = "'22", hjust = 0, colour = "grey40") +
  NULL

figure_place_of_death_home_only %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_place_of_death_home_only", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 18, height = 14, dpi = 300, units = "cm")
```

# Proportion annual deaths by place

Here we'll show 4 lines, one for each place of death, for each year so far.

```{r}
figure_proportion_annual_deaths_by_place <-
  merged_proportions_of_deaths_by_place %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point() +
  geom_line(size = 1) +
  geom_text(data = ~filter(.x, year==2022), aes(x=2022.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  scale_x_continuous(breaks = sort(unique(merged_proportions_of_deaths_by_place$year))) +
  scale_y_continuous(label = function(x) scales::percent(x = x, accuracy = 1)) +
  scale_colour_viridis_d(option="C") +
  # theme_minimal() +
  theme(legend.position = "top", plot.caption = element_text(size = 10, colour = "gray60")) +
  labs(
    x = "Year", y= "Proportion of yearly deaths (%)",
    colour = "",
    subtitle = paste0("* ",only_2022_text),
    caption = paste0(c(source_nrs_text),collapse="\n")
    ) +
  guides(colour = guide_legend(nrow=2)) +  # spread colour legend over two rows
  NULL

figure_proportion_annual_deaths_by_place %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_proportion_annual_deaths_by_place", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 13, height = 10, dpi = 300, units = "cm")
```

# Causes of death over time

## Proportion of causes by place, yearly

This graph shows each of the causes of death divided by place of death - e.g. what proportion of cancer deaths happened at home vs hospital over time?

```{r}
figure_cause_of_death_annual_proportion_by_place <-
  current_and_historical_deaths_by_cause_and_location %>%
  filter(!cause_of_death %in% c("All")) %>%
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home", "Other"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  # mutate(
  #  cause_of_death = factor(cause_of_death %>% str_wrap(., width = 15), levels = order_of_causes_of_death %>% str_wrap(., width = 15)),  # needed to allow converting to a number for plotting the line segments; str_wrap breaks the text up into lines
  # ) %>%
  ggplot(aes(x = year, y = proportion, colour = place_of_death, group = place_of_death)) +
  geom_line() +
  geom_point() +
  facet_rep_wrap(~cause_of_death, repeat.tick.labels = TRUE) +
  scale_colour_viridis_d() +
  scale_y_continuous(labels = function(x) scales::percent(x, accuracy = 1)) +
  labs(
    x = "Year",
    y = "Percentage yearly deaths (%)",
    title = "In 2020 and 2021 the share of deaths at <b style='color:#31688E;'>Home</b> increased for all causes.\nDuring the pandemic, more cancer deaths happened at <b style='color:#31688E;'>Home</b> than in <b style='color:#440154;'>Hospital</b>.",
    subtitle = date_range_text,
    caption = paste0(c("The 2015-2019 number of deaths is the annual mean for the period.",source_nrs_text), collapse = "\n")
  ) +
  theme(
    panel.spacing = unit(1, units = "lines"),
    legend.position = "top", legend.title = element_blank(),
    plot.title = element_textbox_simple()  # to render the html in the title
    ) +
  guides(colour = guide_legend(nrow=2)) +  # spread colour legend over two rows
  NULL

figure_cause_of_death_annual_proportion_by_place %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_cause_of_death_annual_proportion_by_place", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 26, height = 16, dpi = 300, units = "cm")
```

## Deviation from mean

```{r}
figure_place_of_death_deviation_from_mean <-
  merged_deaths_overall %>%
  filter(sex == "all" & age == "all") %>%
  filter(!place_of_death %in% c("All","Other")) %>%
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>%
  mutate(  # subtract mean from all figures to flatten graph
    deaths = deaths - deaths_mean,
    deaths_min = deaths_min - deaths_mean,
    deaths_max = deaths_max - deaths_mean
  ) %>%
  ggplot(aes(x = date_w_c)) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2022)) +
  geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max, fill = "#4477aa"), alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = 0, linetype = "dashed"), colour="grey20") +
  facet_grid(~place_of_death) +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  scale_linetype_manual(values = c("dashed"), labels = "Weekly average (scaled to 0)") +
  scale_fill_manual(values = c("#4477aa"), labels = "Historic weekly minima/maxima") +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = NULL,
       y = "Deaths (difference from weekly average)",
       subtitle = date_range_text,
       caption = paste0(c(death_registrations_text,removed_other_category_text,source_nrs_text), collapse="\n")
       ) +
  annotate(geom = "text", x=ymd("2021-01-28"), y=-250, label = "2021", hjust = 0, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=-250, label = "'22", hjust = 0, colour = "grey40") +
  NULL

figure_place_of_death_deviation_from_mean %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_place_of_death_deviation_from_mean", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 28, height = 14, dpi = 300, units = "cm")
```

@@

```{r}
(figure_home_deaths_over_time_by_hb_4_week_average <-
  merged_deaths_hb %>%
  filter(place_of_death == "Home & other non-institution") %>%
  mutate(ref_area = factor(x = ref_area, levels = order_hb_by_population)) %>%  # arrange hb/la by population size
  pivot_longer(cols = c(ma4w_deaths_all_causes, ma4w_deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2022)) +
  geom_ribbon(aes(ymin = ma4w_deaths_min, ymax = ma4w_deaths_max, fill = "#4477aa"), alpha = 0.5) +
  # geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = ma4w_deaths_mean, linetype = "dashed"), size = 0.2) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  # scale_x_continuous(breaks = seq(0,60,by=10)) +
  facet_wrap(~ref_area, scales = "free_y") +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  scale_linetype_manual(values = c("dashed"), labels = "Weekly average") +
  scale_fill_manual(values = c("#4477aa"), labels = "Historic weekly minima/maxima") +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
    ) +
  labs(x = NULL,
       y = "Deaths (average of past 4 weeks)",
       title = "Deaths at Home*: 4-week average by NHS Health Board",
       subtitle = date_range_text,
       caption = paste0(c(
         "*Home or other non-institutional setting", vertical_line_text, vertical_scale_differs_between_panels_text, source_nrs_text
         ), collapse="\n")
       ) +
  ## annotate historic range
  annotate(geom = "text", x=ymd("2021-01-07"), y=0, label = "2021", hjust = 0, vjust = -1, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=0, label = "'22", hjust = 0, vjust = -1, colour = "grey40") +
  NULL
 )

figure_home_deaths_over_time_by_hb_4_week_average %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_home_deaths_4week_average_over_time_by_hb", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 30, height = 20, dpi = 300, units = "cm")
```
